### 2021-07-21 paper/localpsf.tex
modify notation. Atilde -> Apc. Akernel = Phi

Local translation invariance is illustrated in Figure FIG. If $\mathcal{A}$ were perfectly translation invariant (i.e., if equality held in \eqref{eq:local_translation_invariance} for all $x$, $y$, $h$) then it is straightforward to show that $\mathcal{A}$ would be a convolution operator, and the convolution kernel, $\varphi_x$, would be given by
\begin{equation}
\label{eq:convolution_kernel}
	\varphi_x(y) := \phi_x(y+x).
\end{equation}
The convolution kernel $\varphi_x$ is the result of translating the impulse response $\phi_x$ to re-center it at zero instead of $x$. Locally translation invariant operators act like convolution operators locally, but the convolution kernel, $\varphi_x$, varies as $x$ changes. We therefore approximate $\mathcal{A}$ by a spatially varying weighted sum of convolution operators, $\widetilde{\mathcal{A}} \approx \mathcal{A}$, of the form
\begin{equation}
\label{eq:product_convolution}
	\left(\widetilde{\mathcal{A}}u\right)(v) := \left( v,~ \sum_{i=1}^r \varphi_i \ast \left(w_i \cdot u\right) \right)_{L^2(\Omega)},
\end{equation}
where $f \cdot g$ denotes pointwise multiplication, $f \ast g$ denotes convolution, and $\left(f, g\right)_{L^2(\Omega)}$ denotes the $L^2$ inner product on $\Omega$, for functions $f,g \in L^2(\Omega)$. To avoid nested subscripts, we write 
\begin{equation*}
\varphi_i := \varphi_{x_i} \quad \text{and} \quad \phi_i := \phi_{x_i},
\end{equation*}
with $\varphi_{x_i}$ defined in \eqref{eq:convolution_kernel} and $\phi_{x_i}$ defined in \eqref{eq:impulse_response_defn}, to denote local convolution kernels and impulse responses corresponding to a collection of points $\{x_i\}_{i=1}^r \subset \Omega$ scattered throughout the domain.  The functions $w_i$ are spatially varying weighting functions that are used to interpolate the convolution kernels $\varphi_i$. Our operator approximation method is defined by how we compute the convolution kernels, $\varphi_i$, how we choose the points, $x_i$, what weighting functions, $w_i$, that we use, and how we use the resulting product-convolution approximation to perform linear algebra operations such as solving linear systems. In Section \ref{sec:overview_intro} we summarize the answers to these questions; we will provide detailed answers to these questions in Section \ref{sec:method}.

Approximations of the form \eqref{eq:product_convolution} are known as \emph{product-convolution} approximations, because the action of each term in the sum consists of a pointwise product, followed by a convolution. The more locally translation invariant an operator is (i.e., the smaller the discrepancy between the left hand side and right hand side in \eqref{eq:local_translation_invariance}), the smaller the number of terms that are required in \eqref{eq:product_convolution} to achieve an accurate product-convolution approximation.

@@

Once the $\varphi_i$ and $w_i$ are computed, we convert $\widetilde{\mathcal{A}}$ to hierarchical matrix (H-matrix) format. Let $\widetilde{A}$ denote the integral kernel associated with $\widetilde{\mathcal{A}}$. While $A(y,x)$ is not easily computable, $\widetilde{A}(y,x)$ is given by the formula
\begin{equation}
\label{eq:kernel_entries}
	\widetilde{A}(y,x) = \sum_{i=1}^r \varphi_i(y-x) w_i(x).
\end{equation}
This follows from~\eqref{eq:product_convolution} and the fact that the convolution operator $u \mapsto \varphi_i \ast u$ has $(y,x)$ kernel entry given by $\varphi(y-x)$.
Formula~\eqref{eq:kernel_entries} allows us to construct a H-matrix representation of $\widetilde{\mathcal{A}}$ using the conventional adaptive cross H-matrix construction method, in which one forms low-rank approximations of blocks of the matrix by sampling rows and columns of those blocks \cite{HACA}. Once in H-matrix format, fast H-matrix arithmetic is used to invert $\widetilde{\mathcal{A}}$, or perform other useful matrix operations.

@@

\section{Constructing the approximation, $\widetilde{\mathcal{A}}$}
\label{sec:method}
This section describes how we compute the convolution kernels, $\varphi_i$, and weighting functions, $w_i$, for the product-convolution approximation, $\widetilde{\mathcal{A}} \approx \mathcal{A}$, given in \eqref{eq:product_convolution}. We also describe how we evaluate kernel entries of the approximation, $\widetilde{A}(y,x)$, and how we convert $\widetilde{\mathcal{A}}$ to H-matrix format.

>>>>

Local translation invariance is illustrated in Figure FIG. If $\mathcal{A}$ were perfectly translation invariant (i.e., if equality held in \eqref{eq:local_translation_invariance} for all $x$, $y$, $h$) then it is straightforward to show that $\mathcal{A}$ would be a convolution operator, and the convolution kernel, $\varphi_x$, would be given by
\begin{equation}
\label{eq:convolution_kernel}
	\varphi_x(y) := \phi_x(y+x).
\end{equation}
The convolution kernel $\varphi_x$ is the result of translating the impulse response $\phi_x$ to re-center it at zero instead of $x$. Locally translation invariant operators act like convolution operators locally, but the convolution kernel, $\varphi_x$, varies as $x$ changes. We therefore approximate $\mathcal{A}$ by a spatially varying weighted sum of convolution operators, $\mathcal{A}_\text{pc} \approx \mathcal{A}$, of the form
\begin{equation}
\label{eq:product_convolution}
	\left(\mathcal{A}_\text{pc}~u\right)(v) := \left( v,~ \sum_{i=1}^r \varphi_i \ast \left(w_i \cdot u\right) \right)_{L^2(\Omega)},
\end{equation}
where $f \cdot g$ denotes pointwise multiplication, $f \ast g$ denotes convolution, and $\left(f, g\right)_{L^2(\Omega)}$ denotes the $L^2$ inner product on $\Omega$, for functions $f,g \in L^2(\Omega)$. Approximations of the form \eqref{eq:product_convolution} are known as \emph{product-convolution} approximations, because the action of each term in the sum consists of a pointwise product, followed by a convolution. To avoid nested subscripts, we write 
\begin{equation*}
\varphi_i := \varphi_{x_i} \quad \text{and} \quad \phi_i := \phi_{x_i},
\end{equation*}
with $\varphi_{x_i}$ defined in \eqref{eq:convolution_kernel} and $\phi_{x_i}$ defined in \eqref{eq:impulse_response_defn}, to denote local convolution kernels and impulse responses corresponding to a collection of points $\{x_i\}_{i=1}^r \subset \Omega$ scattered throughout the domain. The functions $w_i$ are spatially varying weighting functions that are used to interpolate the convolution kernels $\varphi_i$. Our operator approximation method is defined by how we compute the convolution kernels, $\varphi_i$, how we choose the points, $x_i$, what weighting functions, $w_i$, that we use, and how we use the resulting product-convolution approximation to perform linear algebra operations such as solving linear systems. In Section \ref{sec:overview_intro} we summarize the answers to these questions; we will provide detailed answers to these questions in Section \ref{sec:method}.

The more locally translation invariant an operator is (i.e., the smaller the discrepancy between the left hand side and right hand side in \eqref{eq:local_translation_invariance}), the smaller the number of terms that are required in \eqref{eq:product_convolution} to achieve an accurate product-convolution approximation.

@@

Once the $\varphi_i$ and $w_i$ are computed, we convert $\mathcal{A}_\text{pc}$ to hierarchical matrix (H-matrix) format. Let $\Phi_\text{pc}$ denote the integral kernel associated with $\mathcal{A}_\text{pc}$. While $\Phi(y,x)$ is not easily computable, $\Phi_\text{pc}(y,x)$ is given by the formula
\begin{equation}
\label{eq:kernel_entries}
	\Phi_\text{pc}(y,x) = \sum_{i=1}^r \varphi_i(y-x) w_i(x).
\end{equation}
This follows from~\eqref{eq:product_convolution} and the fact that the convolution operator $u \mapsto \varphi_i \ast u$ has $(y,x)$ kernel entry given by $\varphi(y-x)$.
Formula~\eqref{eq:kernel_entries} allows us to construct a H-matrix representation of $\mathcal{A}_\text{pc}$ using the conventional adaptive cross H-matrix construction method, in which one forms low-rank approximations of blocks of the matrix by sampling rows and columns of those blocks \cite{HACA}. Once in H-matrix format, fast H-matrix arithmetic is used to invert $\mathcal{A}_\text{pc}$, or perform other useful matrix operations.

Often we are interested in approximating operators $\mathcal{A}$ that are symmetric positive semi-definite. Unfortunately, $\mathcal{A}_\text{pc}$ is, in general, non-symmetric and indefinite due to errors in the product convolution approximation. In this case, we modify the H-matrix representation of $\mathcal{A}_\text{pc}$ to make it positive definite, by symmetrizing it, then applying a specially chosen rational matrix function to the symmetrized H-matrix.

@@

\section{Constructing the product-convolution H-matrix approximation}
\label{sec:method}
This section describes how we compute the convolution kernels, $\varphi_i$, and weighting functions, $w_i$, for the product-convolution approximation, $\mathcal{A}_\text{pc} \approx \mathcal{A}$, given in \eqref{eq:product_convolution}. We also describe how we evaluate kernel entries of the approximation, $\Phi_\text{pc}(y,x)$, how we convert $\mathcal{A}_\text{pc}$ to H-matrix format, and (optionally) how we modify $\mathcal{A}_\text{pc}$ to force it to be symmetric positive semi-definite.
